---
title: "Model Interpretation"
author: "Bodo Winter"
date: "16/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

Load packages:

```{r, message = FALSE}
library(tidyverse)
library(brms)
library(gridExtra)
```

Load language data for plotting:

```{r, message = FALSE}
lang_info <- read_csv('../data/language_info.csv')
field_info <- read_csv('../data/language_info_field.csv')
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

Load models:

```{r}
load('../models/main_mdl.RData')
load('../models/field_animacy_mdl.RData')
load('../models/field_mdl.RData')
```

## Posterior predictive checks:

Perform posterior predictive checks:

```{r, fig.width = 8, fig.height = 6, cache = TRUE}
pp_check(main_mdl, nsamples = 1000)
pp_check(field_mdl, nsamples = 1000)
pp_check(field_animacy_mdl, nsamples = 1000)
```

They all look pretty good! (The black line falls well within the rest of the lines)

## Summarize models, first accuracy-only, without consideration of animacy:

Summarize models, first the main model for the online data:

```{r}
summary(main_mdl)
```

What is the estimated average accuracy?

```{r}
plogis(fixef(main_mdl)[, 1])

# Lower bound of 95% CI:

plogis(fixef(main_mdl)[, 3])

# Upper bound of 95% CI:

plogis(fixef(main_mdl)[, 4])
```

The lower threshold does not include chance level (= 1/6).

Extract posterior samples:

```{r}
main_samples <- posterior_samples(main_mdl)
field_samples <- posterior_samples(field_mdl)
```

What is the posterior probability of this being above the chancel threshold of 1/6?

```{r}
myposts <- main_samples$b_Intercept
sum(myposts > qlogis(1/6)) / length(myposts)
```

100%, not a single posterior sample below the chance threshold of 1/6.

What is the posterior probability of the speaker random effect being larger than the language random effect?

```{r}
sum(main_samples$sd_ID__Intercept > main_samples$sd_Language__Intercept) / nrow(main_samples)
```


Check the same for the field data:

```{r}
summary(field_mdl)
```

What is the estimated average accuracy?

```{r}
plogis(fixef(field_mdl)[, 1])

# Lower bound of 95% CI:

plogis(fixef(field_mdl)[, 3])

# Upper bound of 95% CI:

plogis(fixef(field_mdl)[, 4])
```

The lower threshold does not include chance level (= 1/12).

What is the posterior probability of this being above the chancel threshold of 1/12?

```{r}
myposts <- field_samples$b_Intercept
sum(myposts > qlogis(1/12)) / length(myposts)
```

99%, very few posterior samples below the chance threshold of 1/12.

## Summarize models, with animacy predictors:

Check same for field model:

```{r}
summary(field_animacy_mdl)
```

Check coefficients, first the estimate (times two because sum-coded):

```{r}
round(fixef(field_animacy_mdl)[2, 1] * 2, 2)
```

The standard error also needs to be multiplied by two:

```{r}
round(fixef(field_animacy_mdl)[2, 2] * 2, 2)
```

The odds:

```{r}
round(exp(fixef(field_animacy_mdl)[2, 1] * 2), 2)
```

The posterior probability of the animacy coefficient being above zero:

```{r}
animacy_samples <- posterior_samples(field_animacy_mdl)
sum(animacy_samples$b_Animacy1 > 0) / nrow(animacy_samples)
```

## Plotting the model estimates, online experiment:

Getting the posteriors per language:

```{r}
lang_samples <- select(main_samples, contains('Language'))
head(lang_samples)
```

Get rid of the first, which is across all of them:

```{r}
lang_samples <- lang_samples[, -1]
```

The genus information needs to be added for each language. First, extract the family info:

```{r}
benue_congo <- main_samples$`r_Genus[Benue-Congo,Intercept]`
finno_ugric <- main_samples$`r_Genus[Finno-Ugric,Intercept]`
IE <- main_samples$`r_Genus[IE,Intercept]`
korean <- main_samples$`r_Genus[Korean,Intercept]`
kartvelian <- main_samples$`r_Genus[Kartvelian,Intercept]`
sino_tibetan <- main_samples$`r_Genus[Sino-Tibetan,Intercept]`
tai_kadai <- main_samples$`r_Genus[Tai-Kadai,Intercept]`
turkic <- main_samples$`r_Genus[Turkic,Intercept]`
afro_asiatic <- main_samples$`r_Genus[Afro-Asiatic,Intercept]`
japanese <- main_samples$`r_Genus[Japanese,Intercept]`
```

Extract intercept:

```{r}
int <- main_samples$b_Intercept
```

Add the language family estimates and the intercepts to the respective languages:

```{r}
lang_samples$`r_Language[AL,Intercept]` <- lang_samples$`r_Language[AL,Intercept]` + IE + int
lang_samples$`r_Language[AM,Intercept]` <- lang_samples$`r_Language[AM,Intercept]` + afro_asiatic + int # need to fix
lang_samples$`r_Language[CN,Intercept]` <- lang_samples$`r_Language[CN,Intercept]` + sino_tibetan + int
lang_samples$`r_Language[DE,Intercept]` <- lang_samples$`r_Language[DE,Intercept]` + IE + int
lang_samples$`r_Language[DK,Intercept]` <- lang_samples$`r_Language[DK,Intercept]` + IE + int

lang_samples$`r_Language[EE,Intercept]` <- lang_samples$`r_Language[EE,Intercept]` + finno_ugric + int
lang_samples$`r_Language[EN,Intercept]` <- lang_samples$`r_Language[EN,Intercept]` + IE + int
lang_samples$`r_Language[ES,Intercept]` <- lang_samples$`r_Language[ES,Intercept]` + IE + int
lang_samples$`r_Language[FA,Intercept]` <- lang_samples$`r_Language[FA,Intercept]` + IE + int
lang_samples$`r_Language[FI,Intercept]` <- lang_samples$`r_Language[FI,Intercept]` + finno_ugric + int
  
lang_samples$`r_Language[FR,Intercept]` <- lang_samples$`r_Language[FR,Intercept]` + IE + int
lang_samples$`r_Language[GE,Intercept]` <- lang_samples$`r_Language[GE,Intercept]` + kartvelian + int
lang_samples$`r_Language[GR,Intercept]` <- lang_samples$`r_Language[GR,Intercept]` + IE + int
lang_samples$`r_Language[HU,Intercept]` <- lang_samples$`r_Language[HU,Intercept]` + finno_ugric + int
lang_samples$`r_Language[IT,Intercept]` <- lang_samples$`r_Language[IT,Intercept]` + IE + int
  
lang_samples$`r_Language[JP,Intercept]` <- lang_samples$`r_Language[JP,Intercept]` + japanese + int
lang_samples$`r_Language[KR,Intercept]` <- lang_samples$`r_Language[KR,Intercept]` + korean + int
lang_samples$`r_Language[PL,Intercept]` <- lang_samples$`r_Language[PL,Intercept]` + IE + int
lang_samples$`r_Language[PT,Intercept]` <- lang_samples$`r_Language[PT,Intercept]` + IE + int
lang_samples$`r_Language[RO,Intercept]` <- lang_samples$`r_Language[RO,Intercept]` + IE + int
  
lang_samples$`r_Language[RU,Intercept]` <- lang_samples$`r_Language[RU,Intercept]` + IE + int
lang_samples$`r_Language[SE,Intercept]` <- lang_samples$`r_Language[SE,Intercept]` + IE + int
lang_samples$`r_Language[TH,Intercept]` <- lang_samples$`r_Language[TH,Intercept]` + tai_kadai + int
lang_samples$`r_Language[TR,Intercept]` <- lang_samples$`r_Language[TR,Intercept]` + turkic + int
lang_samples$`r_Language[ZU,Intercept]` <- lang_samples$`r_Language[ZU,Intercept]` + benue_congo + int
```

Compute 95% CI for all of these:

```{r}
lower_CIs <- apply(lang_samples, FUN = function(x) quantile(x, 0.025), MARGIN = 2)
upper_CIs <- apply(lang_samples, FUN = function(x) quantile(x, 0.975), MARGIN = 2)
post_mean <- colMeans(lang_samples)
```

Put those into a table, with logistic transform back to probabilities:

```{r}
lang_intervals <- tibble(Language = str_extract(colnames(lang_samples), '[A-Z]{2,2}'),
                         lower_CI = plogis(lower_CIs),
                         upper_CI = plogis(upper_CIs),
                         M = plogis(post_mean))
```

Merge with name for plotting:

```{r}
lang_intervals$Name <- lang_info[match(lang_intervals$Language, lang_info$Language), ]$Name
```

Change "English UK" label to just "English":

```{r}
lang_intervals[lang_intervals$Language == 'EN', ]$Name <- 'English'
```

"Amharic" is actually "Armenian":

```{r}
lang_intervals[lang_intervals$Name == 'Amharic', ]$Name <- 'Armenian'
```

Change the order of language levels according to the posterior mean:

```{r}
lang_intervals <- arrange(lang_intervals, M)
lang_intervals <- mutate(lang_intervals,
                         Name = factor(Name, levels = lang_intervals$Name))
```

Make a plot with accuracy per language:

```{r, fig.width = 10, fig.height = 6}
lang_plot <- ggplot(data = lang_intervals,
                    mapping = aes(x = Name, y = M,
                                  ymin = lower_CI, ymax = upper_CI)) +
  geom_errorbar(width = 0.3, lwd = 1) +
  geom_point(pch = 22, size = 7, fill = 'red', col = 'black', stroke = 1) +
  geom_hline(yintercept = 1/6, linetype = 2) +
  theme_timo +
  annotate('text', label = 'Chance level', x = 24.5, y = 1/6 + 0.04,
           size = 10, fontface = 'bold', hjust = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1,
                                   face = 'bold', size = 24),
        axis.title.y = element_text(margin = margin(t = 0, r = 50, b = 0, l = 0),
                                    size = 30),
        axis.text.y = element_text(face = 'bold', size = 20),
        plot.title = element_text(face = 'bold', size = 36, hjust = 0,
                                  margin = margin(t = 0, r = 0, b = 20, l = 0))) +
  labs(title = '(a) Estimated accuracy by language',
       x = '', y = 'Posterior probability')
lang_plot
```

Getting the posteriors per meaning:

```{r}
meaning_samples <- select(main_samples, contains('Meaning'))
head(meaning_samples)
```

Get rid of the first, which is across all of them:

```{r}
meaning_samples <- meaning_samples[, -1]
```

Add the intercept to this one as well:

```{r}
for (i in 1:ncol(meaning_samples)) {
  meaning_samples[, i] <- meaning_samples[, i] + int
}
```

Compute 95% CI for all of these:

```{r}
lower_CIs <- apply(meaning_samples, 
                   FUN = function(x) quantile(x, 0.025), MARGIN = 2)
upper_CIs <- apply(meaning_samples,
                   FUN = function(x) quantile(x, 0.975), MARGIN = 2)
post_mean <- colMeans(meaning_samples)
```

Put those into a table, with logistic transform back to probabilities:

```{r}
meaning_intervals <- tibble(Meaning = str_extract(colnames(meaning_samples),
                                                  '\\[[a-z]{2,10}'),
                            lower_CI = plogis(lower_CIs),
                            upper_CI = plogis(upper_CIs),
                            M = plogis(post_mean))
meaning_intervals <- mutate(meaning_intervals,
                            Meaning = str_replace(Meaning, '\\[', ''))
```

Change the order of language levels according to the posterior mean:

```{r}
meaning_intervals <- arrange(meaning_intervals, M)
meaning_intervals <- mutate(meaning_intervals,
                            Meaning = factor(Meaning,
                                             levels = meaning_intervals$Meaning))
```

Make a plot with accuracy per meaning:

```{r, fig.width = 10, fig.height = 6}
meaning_plot <- ggplot(data = meaning_intervals,
                    mapping = aes(x = Meaning, y = M,
                                  ymin = lower_CI, ymax = upper_CI)) +
  geom_errorbar(width = 0.3, lwd = 1) +
  geom_point(pch = 22, size = 7, fill = 'red', col = 'black', stroke = 1) +
  geom_hline(yintercept = 1/6, linetype = 2) +
  theme_timo +
  annotate('text', label = 'Chance level', x = 30.5, y = 1/6 + 0.04,
           size = 10, fontface = 'bold', hjust = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1,
                                   face = 'bold', size = 24),
        axis.title.y = element_text(margin = margin(t = 0, r = 50, b = 0, l = 0),
                                    size = 30),
        axis.text.y = element_text(face = 'bold', size = 20),
        plot.title = element_text(face = 'bold', size = 36, hjust = 0,
                                  margin = margin(t = 0, r = 0, b = 20, l = 0))) +
  labs(title = '(b) Estimated accuracy by meaning',
       x = '', y = 'Posterior probability')
meaning_plot
```

Put both together:

```{r}
my_layout <- rbind(c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(1, 1, 1, 1, 1, 1, 1, 1, NA),
                   c(NA, NA, NA, NA, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2))
multi_plot <- grid.arrange(lang_plot, meaning_plot,
                           layout_matrix = my_layout)
ggsave(plot = multi_plot, '../figures/web_models.pdf',
       width = 24, height = 22)
ggsave(plot = multi_plot, '../figures/web_models.png',
       width = 24, height = 22)
```

## Field experiment:

Getting the posteriors per language:

```{r}
field_lang_samples <- select(field_samples, contains('Language'))
head(field_lang_samples)
```

Get rid of the first, which is across all of them:

```{r}
field_lang_samples <- field_lang_samples[, -1]
```

Add the intercept:

```{r}
for (i in 1:ncol(field_lang_samples)) {
  field_lang_samples[, i] <- field_lang_samples[, i] + field_samples$b_Intercept
}
```

Compute 95% CI for all of these:

```{r}
lower_CIs <- apply(field_lang_samples, FUN = function(x) quantile(x, 0.025), MARGIN = 2)
upper_CIs <- apply(field_lang_samples, FUN = function(x) quantile(x, 0.975), MARGIN = 2)
post_mean <- colMeans(field_lang_samples)
```

Put those into a table, with logistic transform back to probabilities:

```{r}
field_lang_intervals <- tibble(Language = str_extract(colnames(field_lang_samples), '[A-Z]{2,2}'),
                               lower_CI = plogis(lower_CIs),
                               upper_CI = plogis(upper_CIs),
                               M = plogis(post_mean))
```

Add language names for plotting:

```{r}
field_lang_intervals$Name <- character(nrow(field_lang_intervals))
field_lang_intervals[field_lang_intervals$Language == 'DE', ]$Name <- 'German'
field_lang_intervals[field_lang_intervals$Language == 'US', ]$Name <- 'English (US)'
field_lang_intervals[field_lang_intervals$Language == 'BE', ]$Name <- 'English (UK)'
field_lang_intervals[field_lang_intervals$Language == 'BR', ]$Name <- 'Tashlhiyt Berber'
field_lang_intervals[field_lang_intervals$Language == 'PA', ]$Name <- 'Palikur'
field_lang_intervals[field_lang_intervals$Language == 'VA', ]$Name <- 'Daakie'
field_lang_intervals[field_lang_intervals$Language == 'SR', ]$Name <- 'Brazilian Portuguese'
```

Change the order of language levels according to the posterior mean:

```{r}
field_lang_intervals <- arrange(field_lang_intervals, M)
field_lang_intervals <- mutate(field_lang_intervals,
                         Name = factor(Name, levels = field_lang_intervals$Name))
```

Make a plot with accuracy per language:

```{r, fig.width = 10, fig.height = 6}
field_lang_plot <- ggplot(data = field_lang_intervals,
                    mapping = aes(x = Name, y = M, ymin = lower_CI, ymax = upper_CI)) +
  geom_errorbar(width = 0.3, lwd = 1) +
  geom_point(pch = 22, size = 7, fill = 'red',
             col = 'black', stroke = 1) +
  geom_hline(yintercept = 1/12, linetype = 2) +
  theme_timo +
  annotate('text', label = 'Chance level', x = 7.5, y = 1/12 + 0.04,
           size = 10, fontface = 'bold', hjust = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1,
                                   face = 'bold', size = 24),
        axis.title.y = element_text(margin = margin(t = 0, r = 50, b = 0, l = 0),
                                    size = 30),
        axis.text.y = element_text(face = 'bold', size = 20),
        plot.title = element_text(face = 'bold', size = 36, hjust = 0,
                                  margin = margin(t = 0, r = 0, b = 20, l = 0))) +
  labs(title = '(a) Estimated accuracy by language',
       x = '', y = 'Posterior probability')
field_lang_plot
```

Getting the posteriors per meaning:

```{r}
field_meaning_samples <- select(field_samples, contains('Meaning'))
head(field_meaning_samples)
```

Get rid of the first, which is across all of them:

```{r}
field_meaning_samples <- field_meaning_samples[, -1]
```

Add the intercept to all of them:

```{r}
for (i in 1:ncol(field_meaning_samples)) {
  field_meaning_samples[, i] <- field_meaning_samples[, i] + field_samples$b_Intercept
}
```

Compute 95% CI for all of these:

```{r}
lower_CIs <- apply(field_meaning_samples, FUN = function(x) quantile(x, 0.025), MARGIN = 2)
upper_CIs <- apply(field_meaning_samples, FUN = function(x) quantile(x, 0.975), MARGIN = 2)
post_mean <- colMeans(field_meaning_samples)
```

Put those into a table, with logistic transform back to probabilities:

```{r}
field_meaning_intervals <- tibble(Meaning = str_extract(colnames(field_meaning_samples), '\\[[a-z]{2,10}'),
                            lower_CI = plogis(lower_CIs),
                            upper_CI = plogis(upper_CIs),
                            M = plogis(post_mean))
field_meaning_intervals <- mutate(field_meaning_intervals,
                            Meaning = str_replace(Meaning, '\\[', ''))
```

Change the order of language levels according to the posterior mean:

```{r}
field_meaning_intervals <- arrange(field_meaning_intervals, M)
field_meaning_intervals <- mutate(field_meaning_intervals,
                                  Meaning = factor(Meaning, levels = field_meaning_intervals$Meaning))
```

Make a plot with accuracy per meaning:

```{r, fig.width = 10, fig.height = 6}
field_meaning_plot <- ggplot(data = field_meaning_intervals,
                    mapping = aes(x = Meaning, y = M, ymin = lower_CI, ymax = upper_CI)) +
  geom_errorbar(width = 0.3, lwd = 1) +
  geom_point(pch = 22, size = 7, fill = 'red',
             col = 'black', stroke = 1) +
  geom_hline(yintercept = 1/12, linetype = 2) +
  theme_timo +
  annotate('text', label = 'Chance level', x = 12.5, y = 1/12 + 0.04,
           size = 10, fontface = 'bold', hjust = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1,
                                   face = 'bold', size = 24),
        axis.title.y = element_text(margin = margin(t = 0, r = 50, b = 0, l = 0),
                                    size = 30),
        axis.text.y = element_text(face = 'bold', size = 20),
        plot.title = element_text(face = 'bold', size = 36, hjust = 0,
                                  margin = margin(t = 0, r = 0, b = 20, l = 0))) +
  labs(title = '(b) Estimated accuracy by meaning',
       x = '', y = 'Posterior probability')
field_meaning_plot
```

Put both together:

```{r}
my_layout <- rbind(c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(1, 1, 1, 1, 1, 1, NA, NA, NA),
                   c(NA, NA, NA, NA, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2),
                   c(2, 2, 2, 2, 2, 2, 2, 2, 2))
field_multi_plot <- grid.arrange(field_lang_plot, field_meaning_plot,
                           layout_matrix = my_layout)
ggsave(plot = field_multi_plot, '../figures/field_models.pdf',
       width = 20, height = 22)
ggsave(plot = field_multi_plot, '../figures/field_models.png',
       width = 20, height = 22)
```

This completes this analysis.

