---
title: "Preprocessing Data: Field Experiment"
author: "Bodo Winter"
date: "7/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the preprocessing script that cleans the data from the field experiment.

## Load in data

Load packages:

```{r, message = FALSE}
library(tidyverse)
```

Get the data:

```{r, message = FALSE}
# Load:

field <- read_delim('../data/fieldwork/fieldwork-responses-main.csv', delim = ';')
demographics <- read_delim('../data/fieldwork/fieldwork-personal-data.csv', delim = ';')

# Check:

field
demographics

# Check column names:

colnames(field)
```

Fix last column name:

```{r}
colnames(field)[ncol(field)] <- 'PA08'
```

Get rid of the 'dataset' columns:

```{r}
field <- select(field,
                -(`dataset-1`:`dataset-4`))
```

The field response file is wide format. Let's transform this to long format:

```{r}
field <- gather(field, 'ID', 'Choice', -file, -meaning)
```

Get the language info out of the ID column:

```{r}
field <- mutate(field,
                Language = str_extract(ID, '[A-Z]+'))
```

Rename the relevant columns:

```{r}
field <- rename(field,
                UniqueItem = file,
                Meaning = meaning)
```

Add age and gender info:

```{r}
# Rename so that key columns match names:

demographics <- rename(demographics,
                       ID = subject,
                       Age = age,
                       Sex = gender)

# Get only that what is needed:

demographics <- select(demographics,
                       ID, Age, Sex)

# Join:

field <- left_join(field, demographics)

```

Get rid of those that have Choice = NA:

```{r}
field <- filter(field, !is.na(Choice))
```


Exclude participants under 18:

```{r}
field <- filter(field, Age > 18)
```

Get the team info into a separate column:

```{r}
field <- mutate(field,
                Team = str_extract(UniqueItem, '[A-Z]+'))
```


Reorder columns to have participant information, then item info, then response:

```{r}
field <- select(field,
                ID, Language, Age, Sex, Meaning, Team, UniqueItem, Choice)
```

## Clean up the choice column:

Clean up the choice column, first, replacements:

```{r}
field[field$Choice == 'an', ]$Choice <- 'man' # most likely, leads to inaccurate response (biased against our hypothesis)
field[field$Choice == 'chid', ]$Choice <- 'child'
field[field$Choice == 'child,,', ]$Choice <- 'child'
field[field$Choice == 'hild', ]$Choice <- 'child'
field[field$Choice == 'der', ]$Choice <- 'deer'
field[field$Choice == 'dear', ]$Choice <- 'deer'
field[field$Choice == 'deer,,', ]$Choice <- 'deer'
field[field$Choice == 'deer, tiger, man', ]$Choice <- 'deer' # leads to inaccurate response (biased against our hypothesis)
field[field$Choice == 'fie', ]$Choice <- 'fire'
field[field$Choice == 'fire, meat', ]$Choice <- 'fire' # leads to inaccurate response (biased against our hypothesis)
field[field$Choice == 'frit', ]$Choice <- 'fruit'
field[field$Choice == 'fruit,', ]$Choice <- 'fruit'
field[field$Choice == 'fruit, rock', ]$Choice <- 'fruit' # leads to inaccurate response (biased against our hypothesis)
field[field$Choice == 'met', ]$Choice <- 'meat'
field[field$Choice == 'reh', ]$Choice <- 'deer'
field[field$Choice == 'man,,', ]$Choice <- 'man'
field[field$Choice == 'mand', ]$Choice <- 'man'
field[field$Choice == 'mn', ]$Choice <- 'man'
field[field$Choice == 'knife,,', ]$Choice <- 'knife'
field[field$Choice == 'knfe', ]$Choice <- 'knife'
field[field$Choice == 'kife', ]$Choice <- 'knife'
field[field$Choice == 'knif', ]$Choice <- 'knife'
field[field$Choice == 'rock,,', ]$Choice <- 'rock'
field[field$Choice == 'tiger,', ]$Choice <- 'tiger'
field[field$Choice == 'tiger (man)', ]$Choice <- 'tiger'# leads to inaccurate response (biased against our hypothesis)
field[field$Choice == 'tigr', ]$Choice <- 'tiger'
field[field$Choice == 'tiger,,', ]$Choice <- 'tiger'
field[field$Choice == 'snake,,', ]$Choice <- 'snake'
field[field$Choice == 'wat', ]$Choice <- 'water'
field[field$Choice == 'water,', ]$Choice <- 'water'
field[field$Choice == 'water,,', ]$Choice <- 'water'
field[field$Choice == 'wman', ]$Choice <- 'woman'
field[field$Choice == 'woan', ]$Choice <- 'woman'
field[field$Choice == 'soman', ]$Choice <- 'woman'
field[field$Choice == 'woman,,', ]$Choice <- 'woman'
```

Second, exclusions:

```{r}
field <- filter(field, !(Choice %in% c(',,', '0')))
```

Compute accuracy:

```{r}
field <- mutate(field,
                ACC = ifelse(Choice == Meaning, 1, 0))
```

Discard Polish (only one data point):

```{r}
field <- filter(field, Language != 'PL')
```

## Add language names:

Add language name information to the data frame:

```{r}
field$Name <- 'English'
field[field$Language == 'DE', ]$Name <- 'German'
field[field$Language == 'SR', ]$Name <- 'Portuguese'
field[field$Language == 'VA', ]$Name <- 'Daakie'
field[field$Language == 'BR', ]$Name <- 'Tashlhiyt Berber'
field[field$Language == 'PA', ]$Name <- 'Palikur'
```

Create a "Name2" variable that differentiates between British and American English:

```{r}
field$Name2 <- field$Name
field[field$Language == 'US', ]$Name2 <- 'American English'
field[field$Language == 'BE', ]$Name2 <- 'British English'
```

## Merge meaning categories into there:

Load meaning categories:

```{r, message = FALSE}
sem <- read_csv('../data/meaning_categories.csv')
```

Merge this with the accuracy data:

```{r}
field <- left_join(field, sem, by = c('Meaning' = 'meaning'))
```

Create an "Animacy" variable from the "subcategory" column (this lumps "People" and "Animals" together):

```{r}
field <- mutate(field,
                Animacy = ifelse(subcategory == 'inanimate', 'inanimate', 'animate'))
```

## Save output:

Write the file:

```{r, warning = FALSE}
write_csv(field, '../data/field_cleaned.csv')
```









